### 2022-04-22 greta/hopf_mcmc.Rmd
work in progress

```{r}
set.seed(4242)
train_reps <- 10
train_t_max <- 100
test_t_max <- 100
test_reps <- 100

@@

  N[1,] <- c(p$H_init, p$P_init)
  for (t in 1:(t_max-1)) {
    N[t+1,] <- step(N[t,1], N[t,2], eta[t,], t=t, p = p)
  }
  tibble::tibble(t = 1:t_max, H = N[,1], P = N[,2])
}
```

@@

sim |> 
  pivot_longer(c(P,H), values_to="density", names_to="species") |>
  ggplot(aes(t, density, col=species, group=interaction(i,species))) + geom_line(alpha=0.2) +
  geom_vline(aes(xintercept = train_t_max))
```

@@

```{r}
library(greta)
r <- uniform(0, 10)
c <- uniform(0, 10)
Ko <- uniform(0, 50)
delta <- uniform(-1,1)
sigma_H <- uniform(0, 1)
sigma_P <- uniform(0, 1)

@@

posterior_samples <- 
  bind_rows(map(draws, as_tibble)) %>% 
  sample_n(test_reps) |>
  bind_cols(left_off)

posterior_sims <- posterior_samples %>%
  purrr::transpose() %>%
  map_dfr(function(q) simulate(t_max = test_t_max,
                               p = q
                               ) ,.id = "i") |>
  mutate(t = t+train_t_max)
```

```{r}
combined <- 
  bind_rows(
    mutate(train, type="training"), 
    mutate(test,type="observed"),  
    mutate(posterior_sims, type="predicted") |> filter(i %in% 1:3)
    ) |>
  rename("host" = H, "parasite" = P) |>
  pivot_longer(c(host, parasite), values_to="density", names_to="species")

combined |> 
  ggplot(aes(t, density, col=type, group=interaction(i,type))) + 
  geom_line(alpha=0.5) +
  geom_vline(aes(xintercept = train_t_max)) + facet_wrap(~species, ncol=1)
```

@@

>>>>

```{r}
set.seed(4242)
train_reps <- 1
train_t_max <- 100
test_t_max <- 100
test_reps <- 100

@@

  N[1,] <- c(p$H_init, p$P_init)
  for (t in 1:(t_max-1)) {
    N[t+1,] <- step(N[t,1], N[t,2], eta[t,], t=t, p = p)
  }
  tibble::tibble(t = p$t_init + 1:t_max, H = N[,1], P = N[,2])
}
```

# Increasing K

@@

sim |> 
  pivot_longer(c(P,H), values_to="density", names_to="species") |>
  ggplot(aes(t, density, col=species, group=interaction(i,species))) + 
  geom_line() +
  geom_vline(aes(xintercept = train_t_max))
```

@@

```{r}
library(greta)
r <- lognormal(0, 1)
c <- lognormal(0, 1)
Ko <- uniform(0, 50)
delta <- uniform(-1,1)
sigma_H <- lognormal(0, 1)
sigma_P <- lognormal(0, 1)

@@

@@

posterior_samples <- 
  bind_rows(map(draws, as_tibble)) %>% 
  sample_n(test_reps) |>
  mutate(t_init = train_t_max) |>
  bind_cols(left_off)

posterior_sims <- posterior_samples %>%
  purrr::transpose() %>%
  map_dfr(function(q) simulate(t_max = test_t_max,
                               p = q
                               ) ,.id = "i")
```

```{r}
combined <- 
  bind_rows(
    mutate(train, type="training"), 
    mutate(test,type="observed"),  
    mutate(posterior_sims, type="predicted") 
    ) |>
  rename("host" = H, "parasite" = P) |>
  pivot_longer(c(host, parasite), values_to="density", names_to="species")
```

```{r}
combined |> 
  group_by(t,type,species) |> 
  summarise(mean = mean(density), sd = sd(density), .groups = "drop") |> 
 ggplot(aes(t, col=type)) + 
  geom_ribbon(aes(ymin = mean-2*sd, ymax = mean+2*sd, fill=type), alpha=0.5) +
  geom_line(aes(y=mean)) +
  geom_vline(aes(xintercept = train_t_max)) + facet_wrap(~species, ncol=1)
```

```{r}
combined |> 
      filter(i %in% 1:train_reps) |> 
  ggplot(aes(t, density, col=type, group=interaction(i,type))) + 
  geom_line(alpha=0.5) +
  geom_vline(aes(xintercept = train_t_max)) + facet_wrap(~species, ncol=1)
```

# Decreasing K

```{r}
t_max <- train_t_max + test_t_max
 p = list(H_init = 9, P_init = 1,
          r = 0.75, c = 0.1,  Ko = 30, delta = -0.08,
          sigma_H = 2e-2, sigma_P = 2e-2, t_init = 0)
sim <- purrr::map_dfr(1:train_reps, \(i) simulate(t_max=t_max, p = p), .id = "i")
train <- sim |> filter(t <= train_t_max)
test <- sim |> filter(t > train_t_max)
# test <- purrr::map_dfr(1:test_reps, \(i) simulate(t_max=test_t_max), .id = "i")

sim |> 
  pivot_longer(c(P,H), values_to="density", names_to="species") |>
  ggplot(aes(t, density, col=species, group=interaction(i,species))) + geom_line(alpha=1) +
  geom_vline(aes(xintercept = train_t_max))
```

```{r}
library(greta)
r <- lognormal(0, 1)
c <- lognormal(0, 1)
Ko <- uniform(0, 50)
delta <- uniform(-1,1)
sigma_H <- lognormal(0, 1)
sigma_P <- lognormal(0, 1)

```{r}
mmcmc <- memoise::memoise(mcmc, cache = memoise::cache_filesystem("hopf_decreasing_cache"))
bench::bench_time({                 
    draws <- mmcmc(m, 
                 n_samples = 10000, warmup = 5000,
                 chains = 4, verbose = FALSE)
})
```

posterior_samples <- 
  bind_rows(map(draws, as_tibble)) %>% 
  sample_n(test_reps) |>
  mutate(t_init = train_t_max) |>
  bind_cols(left_off)

posterior_sims <- posterior_samples %>%
  purrr::transpose() %>%
  map_dfr(function(q) simulate(t_max = test_t_max,
                               p = q
                               ) ,.id = "i")
```

```{r}
combined <- 
  bind_rows(
    mutate(train, type="training"), 
    mutate(test,type="observed"),  
    mutate(posterior_sims, type="predicted") 
    ) |>
  rename("host" = H, "parasite" = P) |>
  pivot_longer(c(host, parasite), values_to="density", names_to="species")
```

```{r}
combined |> 
  group_by(t,type,species) |> 
  summarise(mean = mean(density), sd = sd(density), .groups = "drop") |> 
 ggplot(aes(t, col=type)) + 
  geom_ribbon(aes(ymin = mean-2*sd, ymax = mean+2*sd, fill=type), alpha=0.5) +
  geom_line(aes(y=mean)) +
  geom_vline(aes(xintercept = train_t_max)) + facet_wrap(~species, ncol=1)
```

```{r}
combined |> 
      filter(i %in% 1:train_reps) |> 
  ggplot(aes(t, density, col=type, group=interaction(i,type))) + 
  geom_line(alpha=0.5) +
  geom_vline(aes(xintercept = train_t_max)) + facet_wrap(~species, ncol=1)
```
